---
title: "finance-equations"
author: "GeoffRussell"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
comma<-function(x) prettyNum(signif(x,digits=2),big.mark=",")
```

## Present value of annuity 

From [Investopedia](https://www.investopedia.com/terms/p/present-value-annuity.asp)
	
$$
PV = PMT Ã— \frac{1 âˆ’ ( \frac{1}{( 1 + ð‘Ÿ )^ð‘›})}{r}
$$

```{r}
PMT=50000
drate=c(0.06)
years=25
pva<-function(PMT,drate,years) {
  PMT * (1 - (1/(1+drate)^years))/drate
}
comma(pva(PMT,drate,years))
comma(PMT*years)
```

```{r annuity}
PMT=17e9
drate=c(0.07)
years=27
pva<-function(PMT,drate,years) {
  PMT * (1 - (1/(1+drate)^years))/drate
}
comma(pva(PMT,drate,years))
comma(PMT*years)
  
```

## Discounted cash flow 

This is supposed to allow you to compare projects taking account of the time value of money.

$$
DCF = \frac{CF_1}{(1+r)^1} + 
\frac{CF_2}{(1+r)^2} + \ldots
\frac{CF_n}{(1+r)^n} 
$$


We can subtract the initial investment $I$ to find the *net present value* (NPV) of the project.


$$
NPV = \frac{CF_1}{(1+r)^1} + 
\frac{CF_2}{(1+r)^2} + \ldots
\frac{CF_n}{(1+r)^n} - I 
$$

Note that in the real world, the sources of uncertainty are both in the $CF$ estimates as well as the discount rate $r$.


## ISP costs 

Using the Frontier Economics report 2024. 

```{r}
cf<-c(15,15,16,16,16,16,16,17,17,18,19,21,22,23,24,25,25,25,25,25,26,27,27,28,28,30,32)
year<-seq(25,51,1)-24
r<-c(0.03,0.07,0.10)
n<-length(cf)
ny<-length(year)
s=0
for (y in year) {
  s=s+cf[y]/((1+r)^y)
}
scosts<-sum(cf)
```

