---
title: "optimalBatterySize"
author: "GeoffRussell"
date: "2025-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
comma<-function(x) prettyNum(signif(x,digits=6),big.mark=",")
options(scipen=9999999)
dopng<-function(name,p,w=2200,h=1300) {
  png(name,width=w,height=h,units="px",res=300,type="cairo-png")
  print(p)
  dev.off()
}
```

Imagine you want to build a baseload system with solar PV and batteries. What is the optimal battery size for a given sized PV farm?

Let's try a few alternatives using some real data. 

We have 55 days from winter 2024 in SA and 19 from Summer 2024 in QLD. The respective capacity factors were 11 and 22 per cent.

```{r}
# Checking facilities in openNem, there was 735 MW available all through this 55 day period
powerSA<-735e6    # as of 26/1/2025
powerQLD<-3830e6    # as of 26/1/2025
dfsummerQLD<-read_csv("openNEMMerge-QLD-30-12-2024-19D.csv") |> select(date,`Solar (Utility) - MW`) |> rename(solar=`Solar (Utility) - MW`)
dfwinterSA<-read_csv("openNEMMerge-SA-10-07-2024-55D.csv") |> select(date,`Solar (Utility) - MW`) |> rename(solar=`Solar (Utility) - MW`)
calcCF<-function(df,power,days) {
  GWh<-sum(df$solar)/12/1e3
  GWh*1e9/(power*24*days)
}
calcCF(dfwinterSA,powerSA,55)
calcCF(dfsummerQLD,powerQLD,19)
```

# Calculating the optimal battery size: baseload

The primary goal is to find the smallest battery size consistent with supplying constant power.  

The constant power has to equal the power times the capacity factor. So if the power was 3 GW and the capacity factor was 33%, then the best we can do would
be 1 GW of constant power. The question then is what sized battery will we need to supply that power?

Algorithm: assume an infinite storage, and known annual capacity factor. Define $CF_p$ as 
the maximum power times the CF. Then when PV output is below $CF_p$, send to grid along with anything in the battery upto but
not exceeding the $CF_p$. If PV output is greater than or equal to $CF_p$, then send $CF_p$ to grid and excess to storage.

```{r}
runCalc<-function(df,power,days,scaleTo,incfp=0,stlimit=1000e9) {
  scaleFac<-1
  if (scaleTo>0) {
    scaleFac<-scaleTo/power
  }
  out<-tibble()
  cfp<-calcCF(df,power,days)*power*scaleFac
  
  efac<-1/12    # factor for converting to energy for 5-minute data
  if (incfp>0) {
    cfp=incfp
    print(paste0("CFp (target): ",comma(cfp/1e6)," MW target energy out per time period: ",comma(cfp*efac/1e6)," MWh"))
  }
  else {
    print(paste0("CFp: ",comma(cfp/1e6)," MW target energy out per time period: ",comma(cfp*efac/1e6)," MWh"))
  }
  eStor<-0 
  eWaste<-0 
  eOut<-0
  tEShort<-0
  nn<-0
  for(i in 1:nrow(df)) {
    p<-df$solar[i]*1e6*scaleFac
    if (p>=cfp) {
      eOut<-eOut+cfp*efac
      eStor<-eStor+(p-cfp)*efac
      stored<-(p-cfp)*efac
      if (eStor>stlimit) {
        eWaste=eWaste+(eStor-stlimit)
        stored<-stored-(eStor-stlimit)
        eStor<-stlimit
      }
      #summaryPVOut=cfp*efac ; summaryStOut=0 ; summaryStored=(p-cfp)*efac ; summaryAvail=eStor
      summaryPVOut=cfp*efac ; summaryStOut=0 ; summaryStored=stored ; summaryAvail=eStor
      case="OK      "
    }
    else {
      eOut<-eOut+p*efac
      summaryPVOut=p*efac 
      shortE<-(cfp-p)*efac
      if (eStor>=shortE) {
        eStor<-eStor-shortE
        summaryStOut=shortE ; summaryStored=0 ; summaryAvail=eStor
        shortE<-0
        case="OK+PV+St"
      }
      else {
        shortE<-shortE-eStor
        summaryStOut=eStor ; summaryStored=0 ; summaryAvail=0
        tEShort<-tEShort+shortE
        eStor<-0
        case="Short  "
      }
    }
#    print(paste0(df$date[i]," case: ",case," PVOut: ",comma(summaryPVOut), " Storage Out:",comma(summaryStOut)," Stored: ",comma(summaryStored)," Avail: ",comma(summaryAvail),"\n"))
    #nn=nn+1
    if (nn>2880) {
      break;
    }
    out<-bind_rows(out,tibble(date=df$date[i],pv=df$solar[i]*scaleFac,pvout=summaryPVOut/efac,stout=summaryStOut/efac,stored=summaryStored,instorage=summaryAvail))
  }
  out 
}
resQLD52G1G<-runCalc(dfsummerQLD,powerQLD,19,5.2e9,1e9)
resQLD52G1Gl19<-runCalc(dfsummerQLD,powerQLD,19,5.2e9,1e9,19e9)
resSA52G1Gl19<-runCalc(dfwinterSA,powerSA,55,5.2e9,1e9,19e9)
```

```{r}
resSA52G1G<-runCalc(dfwinterSA,powerSA,55,5.2e9,1e9)
resSA52G<-runCalc(dfwinterSA,powerSA,55,5.2e9)
resSA1G<-runCalc(dfwinterSA,powerSA,55,1e9)
resSA<-runCalc(dfwinterSA,powerSA,55,0)
resSA38G<-runCalc(dfwinterSA,powerSA,55,3.83e9)
resQLD1G<-runCalc(dfsummerQLD,powerQLD,19,1e9)
resQLD<-runCalc(dfsummerQLD,powerQLD,19,0)
```
```{r}
getPV<-function(.data) {
  .data |> select(date,pv)
}
getPowerOut<-function(.data) {
  .data |> select(date,pvout,stout) |> pivot_longer(cols=c('pvout','stout'),names_to="Output",values_to="MW")
}
getStorage<-function(.data) {
  .data |> select(date,instorage) 
}
doplot<-function(res,pow,scaleTo,tag,blimit=0) {
  scaleFac<-1
  if (scaleTo>0) {
    scaleFac<-scaleTo/pow
  }
  pow<-pow/1e6*scaleFac
  btag="The battery is assumed to have an infinite storage capacity\n"
  if (blimit>0) {
    btag=paste0("The battery capacity is limited to ",comma(blimit),"GWh\n")
  }
  p<-res |> ggplot() + 
    geom_line(aes(x=date,y=pv),color="red")+  
    geom_line(aes(x=date,y=(stout+pvout)/1e6),color="cyan") + 
    labs(x="",y="Megawatts",title=paste0("Smoothing a ",pow," MW solar farm with batteries"),subtitle=paste0("The cyan curve shows the output smoothed by the  use of the battery\n",tag))
  
  q<-res |> getStorage() |> ggplot() + 
    geom_line(aes(x=date,y=instorage/1e6),color="cyan")+
    labs(x="",y="Megawatt-hours",title=paste0("Smoothing a ",pow," MW solar farm with batteries"),
    subtitle=paste0(btag,"The chart shows the energy in the battery over time\n",tag))
  r<-res |> mutate(day=date(date)) |> group_by(day) |> summarise(solar=sum(pvout/12/1e6),battery=sum(stout/12/1e6)) |> pivot_longer(cols=c('solar','battery')) |>
  ggplot() + geom_col(aes(x=day,y=value,fill=name))+
    labs(x="",y="Megawatt-hours/day",title=paste0("Smoothing a ",pow," MW solar farm with batteries"),
    subtitle=paste0("The chart shows the daily energy output\nand the components direct from the sun or via the battery\n",tag))
  
  list(p,q,r)
}

lqld52g1g=doplot(resQLD52G1G,powerQLD,5.2e9,"Underlying data: QLD 2024, openNEM",blimit=0)
lqld52g1g[[1]]
lqld52g1g[[2]]
lqld52g1g[[3]]

lqld52g1gl19=doplot(resQLD52G1Gl19,powerQLD,5.2e9,"Underlying data: QLD 2024, openNEM",blimit=19)
lqld52g1gl19[[1]]
lqld52g1gl19[[2]]
lqld52g1gl19[[3]]

lsa52g1g=doplot(resSA52G1G,powerSA,5.2e9,"Underlying data: SA 2024, openNEM",blimit=0)
lsa52g1g[[1]]
lsa52g1g[[2]]
lsa52g1g[[3]]

lsa52g1gl19=doplot(resSA52G1Gl19,powerSA,5.2e9,"Underlying data: SA 2024, openNEM",blimit=19)
lsa52g1gl19[[1]]
lsa52g1gl19[[2]]
lsa52g1gl19[[3]]
```
```{r}
lsa=doplot(resSA,powerSA,0,"Underlying data: SA 2024, openNEM",blimit=0)
lsa[[1]]
lsa[[2]]
lsa[[3]]
lsa52g=doplot(resSA52G,powerSA,5.2e9,"Underlying data: SA 2024, openNEM",blimit=0)
lsa52g[[1]]
lsa52g[[2]]
lsa52g[[3]]
```
```{r}

lsa1g=doplot(resSA1G,powerSA,1e9,"Underlying data: SA 2024, openNEM",blimit=0)
lsa1g[[1]]
lsa1g[[2]]
lsa1g[[3]]
lsa38g=doplot(resSA38G,powerSA,3.8e9,"Underlying data: SA 2024, openNEM",blimit=0)
lsa38g[[1]]
lsa38g[[2]]
lsa38g[[3]]
lqld=doplot(resQLD,powerQLD,0,"Underlying data: QLD 2024, openNEM",blimit=0)
lqld[[1]]
lqld[[2]]
lqld[[3]]
lqld1g=doplot(resQLD1G,powerQLD,1e9,"Underlying data: QLD 2024, openNEM",blimit=0)
lqld1g[[1]]
lqld1g[[2]]
lqld1g[[3]]
```

```{r}
dopng("QLD-solar-baseload-52l19.png",lqld52g1gl19[[1]])
dopng("QLD-solar-baseload-battery-52l19.png",lqld52g1gl19[[2]])
dopng("QLD-solar-baseload-makeup-52l19.png",lqld52g1gl19[[3]])
dopng("QLD-solar-baseload.png",lqld[[1]])
dopng("QLD-solar-baseload-battery.png",lqld[[2]])
dopng("SA-solar-baseload.png",lsa[[1]])
dopng("SA-solar-baseload-battery.png",lsa[[2]])
dopng("SA-solar-baseload-scaled38.png",lsa38g[[1]])
dopng("SA-solar-baseload-battery-scaled38.png",lsa38g[[2]])
```

```{robsolete}
for(hourlybl in c(1)*utilityMW*3) {
  bl<-hourlybl*24
  for(bmax in c(4)*utilityMW) {
    print(paste("Baseload: ",comma(bl),"  max storage: ",comma(bmax),"\n"))
    bcharge=0
    nn=0
    for(i in 1:nrow(dfs)) {
      e<-dfs$solar[i] * 1e9
      eOut=0
      eReq=bl
      if (e<=bl) {
        eOut=e
        eReq=bl-eOut
        print(paste("Day ",i," all output: ",comma(e)," short: ",comma(eReq),"  bstate:",comma(bcharge/bmax),"\n"))
        e=0
        if (bcharge>0) {
          inc=min(eReq,bcharge)
          eOut=eOut+inc
          eReq=eReq-inc
          print(paste("    battery contributes: ",comma(inc),"\n"))
          bcharge=bcharge-inc
        }
      }
      else {
        eOut=bl
        leftOver=e-bl
        bAvail=bmax-bcharge
        bInc=min(bAvail,leftOver)
        bcharge=bcharge+bInc
        leftOver=leftOver-bcharge
        print(paste("Day ",i," PV meets baseload:",comma(e)," leftOver for charging:",comma(leftOver),"   added charge:",comma(bInc)," state: ",comma(bcharge/bmax),"\n"))
        if (leftOver>0) {
           print(paste("     excess wasted: ",leftOver,"\n"))
        }
      }
      if (nn>30) {
        break;
      }
      nn=nn+1
    }
  }
}
```


