---
title: "pris-scrape"
author: "GeoffRussell"
date: "2024-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(xml2)
library(hash)
reactorfilename<-"prisreactorlist-sep-8-2024.csv"
reactorfilename<-"prisreactorlist-oct-14-2024.csv"
reactorfilename<-"prisreactorlist-jan-15-2025.csv"

detailsfilename<-"prisreactordetails-sep-8-2024.csv"
detailsfilename<-"prisreactordetails-oct-14-2024.csv"
detailsfilename<-"prisreactordetails-jan-15-2025.csv"

outputdetailsfilename<-"prisoutputdetails-jan-15-2025.csv"

dopng<-function(name,p,w=2200,h=1300) {
  png(name,width=w,height=h,units="px",res=300,type="cairo-png")
  print(p)
  dev.off()
}
comma<-function(x) prettyNum(signif(x,digits=2),big.mark=",")
comma3<-function(x) prettyNum(signif(x,digits=3),big.mark=",")

```



## PRIS Reactor database extraction

The older *reactor-build.Rmd* code required all kinds of tricks to make it work. The following is much simpler and doesn't need
a browser daemon (unless rvest is running one invisibly).

```{r pris}
html<-read_html("https://pris.iaea.org/PRIS/CountryStatistics/CountryStatisticsLandingPage.aspx")
```

There seems to be a timing issue if the following code is run with the above line

```{r}
c<-html |> html_element("#sidebar-first") |> html_element("ul") |> html_elements("li") |> html_elements("a") 
country_codes<-tibble(code=(c |> html_attr("href"))) |> mutate(codes=gsub(".*current=(..).*","\\1",c)) |>
  select(codes) |> filter(str_length(codes)==2)
```
  
Now that we have all the country codes, we need to first build a list of reactors for each country and then get the details for each reactor in
the list.

```{r}
## Experimental code to get values from top graphical area
getDataForCountry<-function(code) {
  url<-paste0("https://pris.iaea.org/PRIS/CountryStatistics/CountryDetails.aspx?current=",code)
  df<-read_html(url)
  t<-df |> html_elements("#content .chartValueBox") |> html_text2() 
  t[[4]]
}
# Test this
t<-tibble(totGWe=getDataForCountry("US"),Country="US")
t<-bind_rows(t,tibble(totGWe=getDataForCountry("JP"),Country="JP"))
t<-bind_rows(t,tibble(totGWe=getDataForCountry("GB"),Country="GB"))
t<-bind_rows(t,tibble(totGWe=getDataForCountry("KZ"),Country="KZ"))
t<-bind_rows(t,tibble(totGWe=getDataForCountry("IT"),Country="IT"))
t<-bind_rows(t,tibble(totGWe=getDataForCountry("CA"),Country="CA"))
```

```{r}
getReactorsForCountry<-function(code) {
  url<-paste0("https://pris.iaea.org/PRIS/CountryStatistics/CountryDetails.aspx?current=",code)
  df<-read_html(url)
  t<-df |> html_elements("table") 
  t[[2]] |> html_table() |> mutate(Country=code)
}
# Test this
getReactorsForCountry("JP")
```

The following takes a few minutes and we save the file.

```{r}
reactors<-tibble()
for (c in country_codes$codes) {
  reactors<-bind_rows(reactors,getReactorsForCountry(c))
}
write_csv(reactors,reactorfilename)
```

Now for the function to get the details for each reactor we aren't (yet) interested in the operational record, just construction information.

```{r}
reactors<-read_csv(reactorfilename)
```

Having initialised the hash, we can run the next bit over and over again. It takes about an hour for all the reactors in the database.

```{r}
getReactorDetails<-function(name) {
    name<-str_replace_all(name," ","%20")
    url<-"https://world-nuclear.org/nuclear-reactor-database/details/"
    print(name)
    fullurl<-str_c(url,name)
    l<-safely(read_html,'empty page')(fullurl)
    if (is.null(l$error)) { 
        rdetails<-read_html(fullurl) |> html_elements("table") |> html_table() 
        str(rdetails)
        str(rdetails[[4]])
        dff<-bind_rows(rdetails[[4]])
        str(dff)
        df<-tibble(gwh=dff$`Electricity Supplied (GWh)`) |> filter(gwh!="-") 
        str(df)
        gwh<-sum(as.double(df$gwh))
        str(gwh)
        t<-bind_rows(rdetails[[1]],rdetails[[2]],rdetails[[3]],tibble(X1="Total GWh",X2=as.character(gwh)))
        t |> mutate(reactorName=name) |> rename(attribute=X1,value=X2)
    }
    else {
      print(paste0("Can't find: ",name))
      tibble()
    }
}
getFullReactorDetails<-function(name) {
    thename=name
    name<-str_replace_all(name," ","%20")
    url<-"https://world-nuclear.org/nuclear-reactor-database/details/"
    ccc<-c('Reference Unit Power (MWe)','Annual Time Online (Hours)','Electricity Supplied (GWh)')
    print(name)
    fullurl<-str_c(url,name)
    l<-safely(read_html,'empty page')(fullurl)
    if (is.null(l$error)) { 
        rdetails<-read_html(fullurl) |> html_elements("table") |> html_table() 
        str(rdetails)
        dff<-bind_rows(rdetails[[4]])
        outputdf<-rdetails[[4]] |> pivot_longer(cols=ccc) |> select(Year,name,value) |> mutate(reactor=thename,value=as.numeric(value))
        str(outputdf)
        #str(dff)
        df<-tibble(gwh=dff$`Electricity Supplied (GWh)`) |> filter(gwh!="-") 
        str(df)
        gwh<-sum(as.double(df$gwh))
        str(gwh)
        t<-bind_rows(rdetails[[1]],rdetails[[2]],rdetails[[3]],tibble(X1="Total GWh",X2=as.character(gwh)))
        rv<-t |> mutate(reactorName=name) |> rename(attribute=X1,value=X2)
        return(list(rv,outputdf))
    }
    else {
      print(paste0("Can't find: ",name))
      return (list(tibble(),tibble()))
    }
}
```

```{r test}
# Test the function
testd1<-tibble()
testd2<-tibble()
x<-getFullReactorDetails("BARAKAH-1")
testd1<-bind_rows(testd1,x[[1]])
testd2<-bind_rows(testd2,x[[2]])
x<-getFullReactorDetails("BRUCE-1")
testd1<-bind_rows(testd1,x[[1]])
testd2<-bind_rows(testd2,x[[2]])
```

```{rr skip}
testf<-bind_rows(testf,getReactorDetails("BRUCE-2"))
testf<-bind_rows(testf,getReactorDetails("BRUCE-3"))
testf<-bind_rows(testf,getReactorDetails("BRUCE-4"))
testf<-bind_rows(testf,getReactorDetails("BRUCE-5"))
testf<-bind_rows(testf,getReactorDetails("BRUCE-6"))
testf<-bind_rows(testf,getReactorDetails("BRUCE-7"))
testf<-bind_rows(testf,getReactorDetails("BRUCE-8"))
testf<-bind_rows(testf,getReactorDetails("PICKERING-1"))
testf<-bind_rows(testf,getReactorDetails("PICKERING-2"))
testf<-bind_rows(testf,getReactorDetails("PICKERING-3"))
testf<-bind_rows(testf,getReactorDetails("PICKERING-4"))
testf<-bind_rows(testf,getReactorDetails("PICKERING-5"))
testf<-bind_rows(testf,getReactorDetails("PICKERING-6"))
testf<-bind_rows(testf,getReactorDetails("PICKERING-7"))
testf<-bind_rows(testf,getReactorDetails("PICKERING-8"))
testf<-bind_rows(testf,getReactorDetails("DARLINGTON-1"))
testf<-bind_rows(testf,getReactorDetails("DARLINGTON-2"))
testf<-bind_rows(testf,getReactorDetails("DARLINGTON-3"))
testf<-bind_rows(testf,getReactorDetails("DARLINGTON-4"))
#rdt<-bind_rows(rdt,getReactorDetails("BR-3"))
```

```{r}
# Total GWh for reactors in testf
twh<-as.double(testf |> filter(attribute=="Total GWh") |> summarise(twh=sum(as.double(value))/1000))
subsidy<-20e9
subperkwh=subsidy/(twh*1e12/1e3)
subperkwh
```


## Build the reactor tables

We build two tables, the first `globaldf` has data relating to each reactor as a whole. When it entered the grid, or was shutdown, etc. The
second `yearlydf` contains data about the energy generation in each year of operation.

There are some NAs in this data, e.g., for ARMENIA-1 there is only data for a few years of operation. So NA is not an
indication of the reactor being offline. Some reactors have been shut down and then reopened. Such things complicate the processing

```{r}
yearlydf<-tibble()
globaldf<-tibble()
dthash<-hash()
```

```{r}
nn<-0
for(n in reactors$Name) {
  if (!has.key(n,dthash)) {
    # dthash[n]=getReactorDetails(n)
    l<-getFullReactorDetails(n)
    globaldf<-bind_rows(globaldf,l[[1]])
    yearlydf<-bind_rows(yearlydf,l[[2]])
    dthash[n]=n
    nn<-nn+1
    #if (nn>100) break
  }
}
write_csv(globaldf,detailsfilename)
write_csv(yearlydf,outputdetailsfilename)
```

Now we can collect up all the records from the hash and write to a file for future use

```{robsolete}
keys(dthash)
rdt<-tibble()
for (n in keys(dthash)) {
  if (length(dthash[[n]])==0) {
    print(paste0("No data: ",n))    
  }
  rdt<-bind_rows(rdt,dthash[[n]])
}
rdt |> select(reactorName) |> unique()
write_csv(rdt,detailsfilename)
```
