---
title: "Analyse BP World Energy Statistics 2021"
author: "GeoffRussell"
date: "13/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(readxl)
library(plotly)
```

## Introduction 

This code analyses data in the BP Statistical Review of World Energy spreadsheet; for the year 2021. 

## Read the data 

We read the data and list the sheets in the spreadsheet. The names change from time to time.


```{r data}
bpfile<-"bp-stats-review-2021-all-data.xlsx"
s21<-excel_sheets(bpfile)
print(s21)
```
## Population data

We will be calculating per-capita statistics so need some population data.

We use a down load of the 2020 World Bank Population data. It's latest year is 2019, so we use a polynomial model
to extrapolate for years after that; out to 2040. It's not the world's best population modelling, but good
enough for our purposes.

One wrinkle is that the World Bank country names don't match the BP country names. We throw in ad-hoc renames as required.

```{r pop}
dfpop<-read_csv("API_SP.POP.TOTL_DS2_en_csv_v2_2252106.csv",skip=4)
dfpop$`Country Name` <- gsub("United States","US",dfpop$`Country Name`)
#-----------------------------------------------
# getpop returns a closure which is vectorised, we extrapolate past the last data point, 2019
#-----------------------------------------------
getpop<-function(country) {
  f<-function(yr) {
    rp<-dfpop %>% filter(`Country Name`==country) %>% pivot_longer(names_to="year",cols="1960":"2019") %>% select(year,value)
    m1<-lm(value~poly(as.numeric(year),2),rp)
    df<-tibble(year=seq(1960,2040),Country=country)
    rowp<-df %>% add_predictions(m1) %>% filter(year %in% yr) %>% rename(Year=year,Popn=pred)
  }
}
getPopForList<-function(countries,years=seq(1975,2021)) {
  df<-tibble()
  for(c in countries) {
    df<-bind_rows(df,getpop(c)(years))
  }
  df 
}
# Here's a test call
dfcs<-getPopForList(c("Australia","Germany","France","US"))
```

## Compare historical energy roll out speeds

The first chart in this project is to compare renewable and nuclear roll out speeds. The most rapid
nuclear rollout various between countries, but it is roughly 1975-1990 in France. So we get
data between 1975 and 2020 for various technologies and plot them.


```{r}
# plenty of juggling due to how read_excel names its columns
f<-function(.data,name,startyear="1975",endyear="2020...57") {
  rename(.data,Country=`Terawatt-hours`) %>% mutate(Type=name) %>% filter(!is.na(Country)) %>% 
    filter(!grepl("^Total ",Country)) %>% filter(!grepl("^Other ",Country)) %>% slice(1:84) %>% 
    select(Country,'Type',startyear:endyear) %>% rename('2020'=endyear) %>% 
    pivot_longer(startyear:'2020',names_to="yy",values_to="TWh") %>% filter(!is.na(TWh))
}
getSheet<-function(sheetname,cols,label,startyear="1975") {
    endyear<-paste0("2020...",toString(cols-3))
    print(endyear)
    read_excel(bpfile,sheet=sheetname,skip=2,na='n/a',col_types=c('text',rep('numeric',cols-1))) %>%
            f(label,startyear=startyear,endyear=endyear) 
}
twh<-bind_rows(
    getSheet("Elec Gen from Gas",40,'Gas',startyear="1985"),
    getSheet("Elec Gen from Coal",40,'Coal',startyear="1985"),
    getSheet("Elec Gen from Oil",40,'Oil',startyear="1985"),
    getSheet("Solar Generation - TWh",60,'Solar'),
    getSheet("Wind Generation - TWh",60,'Wind'),
    getSheet("Hydro Generation - TWh",60,'Hydro'),
    getSheet("Nuclear Generation - TWh",60,'Nuclear'),
    getSheet("Geo Biomass Other - TWh",60,'Biomass/Geo')
  ) %>% mutate(Year=as.numeric(yy)) %>% select(-yy) 
getCountryData<-function(.data,countries) {
    dfc<-getPopForList(countries)
    .data %>% filter(Country %in% countries) %>% left_join(dfc)
}
#---------------------------------------------------------
# twh has data for all countries and getCountryData will filter out
# a selected set and add in the population data
#---------------------------------------------------------
ctwh<-twh %>% getCountryData(c("Australia","US","France","Germany","Belgium","Sweden"))
totaltwh <- ctwh %>% group_by(Country,Year) %>% summarise(total=sum(TWh)) %>% 
  select(Country,Year,total) %>% left_join(ctwh %>% select(Country,Year,Popn) %>% unique()) 

p<-ctwh %>% filter(!Type %in% c("Gas","Coal","Oil")) %>% ggplot(aes(x=Year,y=TWh*1e12/Popn/1e6))+
  geom_col(aes(fill=Type))+geom_line(aes(x=Year,y=total*1e12/Popn/1e6,color="black"),data=totaltwh)+
  guides(fill=guide_legend("Show only"),color=guide_legend("Total"))+
  scale_color_manual(values=c("black"),)+
  facet_wrap(~Country)+
  labs(y="Megawatt-hours per capita from clean(ish) energy sources",
       title="Electricity production by source (Data: BP Stats 2021) ")+
  theme(axis.title=element_text(size=7),axis.text.x=element_text(size=7))
p
#--------------------------------------------------------------------------------------
# now make a png for posting
#--------------------------------------------------------------------------------------
png("cleanish-megawatt-hours-per-person.png",width=1800,height=800,units="px",res=300,type="cairo-png")
p
dev.off()
```


