---
title: "Analyse Energy Institute World Energy Statistics"
author: "GeoffRussell"
date: "13/10/2021"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(readxl)
library(plotly)
library(ggrepel)
library(geomtextpath)
library(zoo)
comma<-function(x) prettyNum(signif(x,digits=4),big.mark=",")
comma3<-function(x) prettyNum(signif(x,digits=3),big.mark=",")
comma2<-function(x) prettyNum(signif(x,digits=2),big.mark=",")
dopng<-function(name,p,w=2200,h=1300) {
  png(name,width=w,height=h,units="px",res=300,type="cairo-png")
  print(p)
  dev.off()
}
```

## Introduction 

This code originally analysed data in the BP Statistical Review of World Energy spreadsheet;  for the year 2021. 

The latest update is to analyse the Energy Institute World Energy Statistics from 2023.


## Read the data 

The names of the sheets change from time to time, so when each year comes out, the code may have to be adjusted.


```{r data}
theyear<-"2021D"
theyear<-"2023D"
theyearonly<-"2021"
bpfile<-"bp-stats-review-2021-all-data.xlsx"
#--------------------------------------
theyearonly<-"2023"
theyear<-"2023D"
bpfile<-"Statistical_Review_of_World_Energy_2023.xlsx"
esheetname="Electricity Generation"
peccolname="Primary Energy Consumption"
colnr=41
colnrpec=61
colpfx="2022..."
colpfxend="2022...59"
cfilend="2022...39"
csyear="2022"
gassheet="Elec Gen from Gas"
coalsheet="Elec Gen from Coal"
oilsheet="Elec Gen from Oil"

#--------------------------------------
theyearonly<-"2024"
theyear<-"2024"
bpfile<-"Statistical_Review_of_World_Energy_2024.xlsx"
esheetname="Electricity Generation - TWh"
peccolname="Primary energy cons - EJ"
colnr=42
colnrpec=62
colpfx="2023..."
colpfxend="2023...60"
csyear="2023"
cfilend="2023...40"
gassheet="Gas inputs - Elec generation"
coalsheet="Coal inputs - Elec generation "
oilsheet="Oil inputs - Elec generation "
lngexsheet="Gas - LNG exports bcm"
#--------------------------------------
s21<-excel_sheets(bpfile)
print(s21)
```
## Population data

We will be calculating per-capita statistics so need some population data.

We use a download of the 2022 United Nations global propulation projections.

One wrinkle is that the country names don't match the BP country names. So we change names of things we may need.


```{r pop}
#dfpop<-read_csv("API_SP.POP.TOTL_DS2_en_csv_v2_2252106.csv",skip=4,show_col_types=FALSE) %>% rename(Country=`Country Name`)

dfpop<-read_csv("WPP2022_extract_medium.csv") %>% rename(Country=`Location`,Year=Time,Popn=TPopulation1Jan) %>% mutate(Popn=Popn*1000)
dfpop$Country <- gsub("United States of America","US",dfpop$Country)
dfpop$Country <- gsub("Republic of Korea","South Korea",dfpop$Country)
```



```{r}
#------------------------------------------------------------------------------------------------
# getpop returns a closure which is vectorised, we extrapolate past the last data point, 2019
# we unname pred because we don't need names and they confuse me when looking at structures!
#------------------------------------------------------------------------------------------------
#----------------------------------------------------------------------
# Remove any countries in twh for which we don't have population data
#---------------------------------------------------------------------
removeMissing<-function(.data,df) {
  semi_join(.data,df %>% select(Country),by=c("Country")) 
}
getpop<-function(country) {
  f<-function(yr) {
    rp<-dfpop %>% filter(Country==country) %>% filter(Year %in% yr) 
# We don't need this any more, the UN data has projections out to 2100
#    rp<-dfpop %>% filter(Country==country) %>% pivot_longer(names_to="year",cols=1960:as.numeric(theyearonly)) 
#             %>% select(year,value)
#    m1<-lm(value~poly(as.numeric(year),2),rp)
#    df<-tibble(year=seq(1960,2040),Country=country)
#    rowp<-df %>% add_predictions(m1) %>% filter(year %in% yr) %>% mutate(Year=year,Popn=unname(pred)) %>% select(-year,-pred) 
  }
}
getPopForList<-function(countries,years=seq(1965,as.numeric(theyearonly)-1)) {
  df<-tibble()
  for(c in countries) {
    df<-bind_rows(df,getpop(c)(years))
  }
  df$Popn<-unname(df$Popn)
  df 
}
getTotalTWhForList<-function(countries,years=seq(1975,as.numeric(theyearonly)-1)) {
  df<-tibble()
  for(c in countries) {
    df<-bind_rows(df,(countryTWh %>% filter(Country==c)))
  }
  df$totalTWh<-unname(df$totalTWh)
  df 
}
# Here's a test call
tgetpop<-getpop("Australia")(1975:2000)
tgetpop50<-getpop("Australia")(1975:2050)
tgetPopForList<-getPopForList(c("Australia","Germany","France","US","World","Europe"))
mergePopData<-function(.data,countries) {
    dfc<-getPopForList(countries)
    .data %>% filter(Country %in% countries) %>% left_join(dfc)
}
mergeTotalTWhData<-function(.data,countries) {
    dfe<-getTotalTWhForList(countries)
    .data %>% filter(Country %in% countries) %>% left_join(dfe)
}
```



## Functions used to read and filter sheets 

The process is conceptually simple, but the details are tedius. We measure the change in energy production over some
given number of years for each technology or category of technologies and then rank countries. 

The code is complicated by changes in the column names as the data source moved from BP to the Energy Institute and
also by the EI just arbitrarily changing column names in 2024.


```{r warning=FALSE}
# plenty of data juggling due to how read_excel names its columns
# this code will have to change in 2022!
countryFilter<-function(.data,name,startyear="1975",endyear=cfilend) {
  rename(.data,Country=`Terawatt-hours`) %>% mutate(Type=name) %>% filter(!is.na(Country)) %>% 
    filter(!grepl("^Total ",Country)) %>% filter(!grepl("^Other ",Country)) %>% slice(1:84) %>% 
    select(Country,'Type',startyear:endyear) %>% rename(!!csyear:=endyear) %>% 
    pivot_longer(startyear:!!csyear,names_to="yy",values_to="TWh") %>% filter(!is.na(TWh))
}
energyFilter<-function(.data,name,startyear="1965",endyear=colpfxend) {
  dd<-rename(.data,Country=`Exajoules`) %>% mutate(Type=name) %>% filter(!is.na(Country)) 
    #filter(!grepl("^Total (A|M|C|S|N)",Country,perl=TRUE)) %>% 
  write_csv(dd,"dd1.csv")
  ee<-dd %>% filter(!grepl("^Other ",Country)) %>% slice(1:104) %>% 
    filter(!grepl("of which",Country)) %>%
    select(Country,'Type',startyear:endyear) %>% rename(!!csyear:=endyear) %>% 
    pivot_longer(startyear:!!csyear,names_to="Year",values_to="Exj") %>% filter(!is.na(Exj))
  ee$Country <- gsub("Total ","",ee$Country)
  ee$Country <- gsub("Union #","Union",ee$Country)
  write_csv(ee,"dd2.csv")
  ee
}
worldTotalFilter<-function(.data,name,startyear="1975",endyear=cfilend) {
  rename(.data,Country=`Terawatt-hours`) %>% mutate(Type=name) %>% filter(!is.na(Country)) %>% 
    filter(grepl("^Total World",Country))  %>% 
    select(Country,'Type',startyear:endyear) %>% rename(!!csyear:=endyear) %>% 
    pivot_longer(startyear:!!csyear,names_to="yy",values_to="TWh") %>% filter(!is.na(TWh))
}
# don't know how to do this ... sub china
worldTotalFilterSubChina<-function(.data,name,startyear="1975",endyear=cfilend) {
  rename(.data,Country=`Terawatt-hours`) %>% mutate(Type=name) %>% filter(!is.na(Country)) %>% 
    filter(Country %in% c("Total World","China")) %>% 
    select(Country,'Type',startyear:endyear) %>% rename(!!csyear:=endyear) %>% 
    pivot_longer(startyear:!!csyear,names_to="yy",values_to="TWh") %>% filter(!is.na(TWh))
}
#xxx<-data.frame()
getSheet<-function(sheetname,cols,label,startyear="1975",filter=countryFilter) {
    endyear<-paste0(colpfx,toString(cols-2))
    print(endyear)
    #xxx<<-read_excel(bpfile,sheet=sheetname,skip=2,na='n/a',col_types=c('text',rep('numeric',cols-1),'skip')) 
    read_excel(bpfile,sheet=sheetname,skip=2,na='n/a',col_types=c('text',rep('numeric',cols-1),'skip')) %>%
            filter(label,startyear=startyear,endyear=endyear) %>%
            mutate(Year=as.numeric(yy)) %>% select(-yy) 
}
tst<-read_excel(bpfile,sheet=esheetname,skip=2,na='n/a',col_types=c('text',rep('numeric',colnr-1),'skip'))
electricityUsage<-getSheet(esheetname,colnr,'Electricity',startyear="1985")
```

## LNG



```{r}
options(scipen=100000)
ePerTonneLNG=14.9e6   # Wh/tonne thermal
ePerTonneUth=924e12/7679   # Wh/tonne thermal 
ePerTonneUe=924e12/7679/3   # Wh/tonne thermal 

lng<-read_excel(bpfile,sheet=lngexsheet,skip=2,na='n/a',col_types=c('text',rep('numeric',24),'skip','skip','skip')) %>% 
  rename("Country"="Billion cubic metres") %>%
    filter(!grepl("^Total ",Country)) %>% filter(!grepl("^Other ",Country)) %>% 
    slice(1:18)  
lnglong<-lng  %>% arrange(desc(`2023`)) %>% slice(1:8) %>%
  pivot_longer(cols=c('2000':'2023'),names_to="Year",values_to="BCM") %>% mutate('mt'=BCM*0.73,tUth=mt*1e6*(ePerTonneLNG/ePerTonneUth))
p<-lnglong |> ggplot() + geom_line(aes(x=Year,y=mt,color=Country,group=Country))+
  labs(x="",y="Million tonnes of LNG",title="Global LNG export growth since y2k\nData: Energy Institute 2024")+ 
  scale_y_continuous(sec.axis=sec_axis(~. *9708/78.3,name="Thermal equivalent tonnes of uranium"))+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p
dopng("Global-lng-since-y2k-2023.png",p)

```

## Electricity usage over time 

```{r}
popForUS<-getPopForList(c("US"))
electricityUsagePop <- left_join(electricityUsage,popForUS)
electricityUsagePop |> filter(Country=="US") |> ggplot() + 
  geom_line(aes(x=Year,y=TWh),color="blue") + 
  geom_line(aes(x=Year,y=TWh*1e12/Popn/1e6*300),color="red") + 
  scale_y_continuous(sec.axis=sec_axis(~./300,name="Megawatt hours per person"))+
  labs(x="",y="Terawatt hours",title="US Annual Electricity Usage\nTotal and Per person\nData: Energy Institute 2024")+
  theme(axis.text.y.right=element_text(color="red"),
    axis.title.y.right=element_text(color="red"))

electricityUsagePop |> filter(Country=="US") |> ggplot() +  geom_line(aes(x=Year,y=TWh),color="blue") + 
  labs(x="",y="Terawatt hours",title="US Annual Electricity Usage\nData: Energy Institute 2024")

electricityUsagePop |> filter(Country=="US") |> ggplot() + 
  geom_line(aes(x=Year,y=TWh*1e12/Popn/1e6),color="red")  +
  labs(x="",y="MWh per person",title="US Annual Electricity Usage\nData: Energy Institute 2024")

```

## Oil production 


```{r}
ePerTonne=11.63e6 # Wh of energy in tonne of oil
#eEqPerTonne=11.63e6/3 # mWh of energy in tonne of oil (assuming oil used to generate electricity)

oFilter<-function(.data,name,startyear="2013",endyear=csyear) {
  rename(.data,Country=name) %>% filter(!is.na(Country)) %>%  filter(row_number()<=which(Country=="Total World"))
}
getFFSheet<-function(sheetname,cols=24,skcols=3,skip=2,startyear="1970",endyear=csyear,label="Terawatt-hours",filter=oFilter) {
    read_excel(bpfile,sheet=sheetname,skip=skip,na='n/a',col_types=c('text',rep('numeric',cols),rep('skip',skcols))) %>%
            filter(label,startyear=startyear,endyear=endyear)
}

oilp<-getFFSheet("Oil Production - tonnes",cols=colnr+17,startyear="1965",skcols=3,label="Million tonnes") %>%
  filter(!grepl("Total",Country))
oilpLong<-oilp %>% pivot_longer(cols=`1965`:!!csyear,names_to="Year",values_to="MT") %>% mutate(TWhEq=MT*ePerTonne*1e6/1e12) %>% mutate(Year=as.numeric(Year)) %>% filter(!(Country=="Russian Federation"& Year<1985))


oilpWorld <- oilpLong %>% group_by(Year) %>% summarise(MT=sum(MT),TWhEq=sum(TWhEq)) %>% mutate(Country="World") 

p<-oilpWorld %>% ggplot()+geom_line(aes(x=Year,y=TWhEq))+labs(x="",y="TWh (energy)",title="Global Oil Production")

p
png(paste0("global-oil-",csyear,".png"),width=1800,height=1200,units="px",res=300,type="cairo-png")
p
dev.off()

bigc<-(oilpLong %>% filter(Year==csyear,MT>200))$Country

popForCountry<-getPopForList(bigc)
p<-oilpLong %>% filter(Country %in% bigc) %>% left_join(popForCountry) %>% mutate(mWhPerPerson=TWhEq*1e12/Popn/1e6) %>%
  ggplot(aes(x=Year,y=TWhEq,color=Country))+geom_line()+geom_textline(aes(label=Country),vjust=-0.2,hjust=0.7,text_smoothing=50)+
  annotate("text",x=2010,y=3,label=paste0("Data: Energy Institute ",theyear))+
  labs(x="",y="TWh (energy equivalent)",
    title="Major Oil Producers")+theme(legend.position="none")
p
png(paste0("major-oil-",csyear,".png"),width=1800,height=1200,units="px",res=300,type="cairo-png")
p
dev.off()
```
## Gas production 

```{r}
ejToTWh=277.78
gasp<-getFFSheet("Gas Production - EJ",cols=colnr+12,startyear="1970",skcols=3,label="Exajoules") %>%
  filter(!grepl("Total",Country))
gaspLong<-gasp %>% pivot_longer(cols=`1970`:!!csyear,names_to="Year",values_to="EJ") %>% mutate(TWhEq=EJ*ejToTWh) %>% mutate(Year=as.numeric(Year)) %>% filter(!(Country=="Russian Federation"& Year<1985))


gaspWorld <- gaspLong %>% group_by(Year) %>% summarise(TWhEq=sum(TWhEq)) %>% mutate(Country="World") 

bigc<-(gaspLong %>% filter(Year==csyear,EJ>5))$Country

p<-gaspLong %>% filter(Country %in% bigc) %>% left_join(popForCountry) %>% mutate(mWhPerPerson=TWhEq*1e12/Popn/1e6) %>%
  ggplot(aes(x=Year,y=TWhEq*3,color=Country))+geom_line()+geom_textline(aes(label=Country),vjust=-0.2,hjust=0.7,text_smoothing=50)+
  annotate("text",x=2010,y=500,label=paste0("Data: Energy Institute ",theyear))+
  labs(x="",y="TWh (energy equivalent)",
    title="Major Gas Producers")+theme(legend.position="none")
p
png(paste0("major-gas-",csyear,".png"),width=1800,height=1200,units="px",res=300,type="cairo-png")
p
dev.off()

p<-gaspWorld %>% 
  ggplot(aes(x=Year,y=TWhEq))+geom_line()+labs(x="",y="TWh (energy) per person",
    title="Global Gas Production")+
  annotate("text",x=2010,y=5.8,label=paste0("Data: Energy Institute ",theyear))
p
png(paste0("global-gas-",csyear,".png"),width=1800,height=1200,units="px",res=300,type="cairo-png")
p
dev.off()

# We use the energy equivalent 
p<-gaspWorld %>% left_join(tgetPopForList) %>% mutate(mWhPerPerson=TWhEq*1e12/Popn/1e6) %>%
  ggplot(aes(x=Year,y=mWhPerPerson))+geom_line()+labs(x="",y="mWh (energy) per person",
    title="Per Person Global Gas Production")+
  annotate("text",x=2010,y=5.8,label=paste0("Data: Energy Institute ",theyear))
p
png(paste0("per-person-global-gas-",csyear,".png"),width=1800,height=1200,units="px",res=300,type="cairo-png")
p
dev.off()
```


```{r}

p<-oilpWorld %>% left_join(tgetPopForList) %>% mutate(mWhPerPerson=TWhEq*1e12/Popn/1e6) %>%
  ggplot(aes(x=Year,y=mWhPerPerson))+geom_line()+labs(x="",y="mWh (energy) per person",
    title="Per Person Global Oil Production")+
  annotate("text",x=2010,y=5.8,label=paste0("Data: Energy Institute ",theyear))
p
png(paste0("per-person-global-oil-",csyear,".png"),width=1800,height=1200,units="px",res=300,type="cairo-png")
p
dev.off()

oilc<-getFFSheet("Oil Consumption - Tonnes",cols=colnr+17,startyear="1965",skcols=3,label="Million tonnes") %>%
  filter(!grepl("Total",Country))

popForCountry<-getPopForList(c(oilc$Country,"World"))


oilcLong<-oilc %>% pivot_longer(cols=`1965`:!!csyear,names_to="Year",values_to="MT") %>% mutate(TWhEq=MT*ePerTonne*1e6/1e12) %>%
  mutate(Year=as.numeric(Year)) %>% left_join(popForCountry) %>% mutate(MWhPerCap=TWhEq*1e12/Popn/1e6)

oilcWorld <- oilcLong %>% group_by(Year) %>% summarise(MT=sum(MT),TWhEq=sum(TWhEq)) %>% mutate(Country="World") %>%
  left_join(tgetPopForList) %>% mutate(Year=as.numeric(Year)) %>% left_join(popForCountry) %>%
  mutate(MWhPerCap=TWhEq*1e12/Popn/1e6)

oilcall<-bind_rows(oilcWorld,oilcLong)

  
oilcall %>%  filter(Country %in% c("World","Australia","US","France","Germany","South Africa","China")) %>%
  ggplot(aes(x=Year,y=MWhPerCap,color=Country))+geom_line()+labs(x="",y="mWh (energy) per person",
    title="Per Person Global Oil Consumption")


#t<-oilcLong %>% left_join(popForCountry)
  
```

## Wind growth rates 

```{r wind}
wFilter<-function(.data,name,startyear="2013",endyear=csyear) {
  rename(.data,Country=name) %>% filter(!is.na(Country)) %>%  filter(row_number()<=which(Country=="Total World"))
}
getWindSheet<-function(sheetname,cols=27,skcols=3,skip=2,startyear="1997",endyear=csyear,label="Terawatt-hours",filter=wFilter) {
    read_excel(bpfile,sheet=sheetname,skip=skip,na='n/a',col_types=c('text',rep('numeric',cols),rep('skip',skcols))) %>%
            filter(label,startyear=startyear,endyear=endyear)
}
windCap<-getWindSheet("Wind Installed Capacity",label="Megawatts",skip=3) %>% filter(!grepl("Total",Country))
windLong<-windCap %>% pivot_longer(cols=`1997`:!!csyear,names_to="Year",values_to="MW") 

windWorld <- windLong %>% group_by(Year) %>% summarise(MW=sum(MW)) %>% mutate(Country="World")
```

## Nuclear growth rates 

There is no "Nuclear installed capacity" sheet.


## Solar growth rates 
```{r solargrowth}
sFilter<-function(.data,name,startyear="2013",endyear=csyear) {
  rename(.data,Country=name) %>% filter(!is.na(Country)) %>%  filter(row_number()<=which(Country=="Total World"))
}
getSolarSheet<-function(sheetname,cols=24,skcols=3,skip=2,startyear="2000",endyear=csyear,label="Terawatt-hours",filter=sFilter) {
    read_excel(bpfile,sheet=sheetname,skip=skip,na='n/a',col_types=c('text',rep('numeric',cols),rep('skip',skcols))) %>%
            filter(label,startyear=startyear,endyear=endyear)
}
solarCap<-getSolarSheet("Solar Installed Capacity",label="Megawatts",skip=3) %>% filter(!grepl("Total",Country))

solarLong<-solarCap %>% pivot_longer(cols=`2000`:!!csyear,names_to="Year",values_to="MW") 

solarWorld <- solarLong %>% group_by(Year) %>% summarise(MW=sum(MW)) %>% mutate(Country="World")

tmp<-bind_rows(solarLong,solarWorld)
chinaRest<-tmp %>% pivot_wider(id_cols=c(`Year`),names_from=`Country`,values_from=`MW`) %>% mutate(Rest=World-China) %>% 
  select(Year,China,Rest) %>% pivot_longer(cols=`China`:`Rest`,names_to="Region",values_to="MW") 

chinaRest %>% filter(as.numeric(Year)>2007) %>% 
  ggplot() + geom_col(aes(x=Year,y=MW/1000,fill=`Region`),position="fill") + coord_flip() +
  labs(x="",y="Percent",title="China's percentage of the solar capacity installation")

solarLong %>% filter(Country == "Australia") %>% 
  ggplot() + geom_col(aes(x=Year,y=MW/1000)) + coord_flip() +
  labs(x="",y="Percent",title="Australia's solar capacity growth")



```

## Battery and metal information is now included in the Energy Institute data 

```{r storage}
cFilter<-function(.data,name,startyear="2013",endyear=csyear) {
  rename(.data,Country=name) %>% filter(!is.na(Country)) %>%  filter(row_number()<=which(Country=="Total World"))
  
#    filter(!grepl("^Total ",Country)) %>% filter(!grepl("^Other ",Country)) %>% 
#    select(Country,startyear:!!endyear) %>% 
#    pivot_longer(startyear:!!endyear,names_to="yy",values_to="GW") %>% filter(!is.na(GW))
}
getBessSheet<-function(sheetname,cols=12,skcols=3,startyear="2013",endyear=csyear,label="Installed Capacity (Gigawatts)",filter=cFilter) {
    read_excel(bpfile,sheet=sheetname,skip=2,na='n/a',col_types=c('text',rep('numeric',cols-1),rep('skip',skcols))) %>%
            filter(label,startyear=startyear,endyear=endyear)
}
bess<-getBessSheet("Grid Scale BESS Capacity")

```


```{r}

getEnergy<-function(sheetname,cols,startyear="1965",filter=energyFilter) {
    endyear<-paste0(colpfx,toString(cols-2))
    read_excel(bpfile,sheet=sheetname,skip=2,na='n/a',col_types=c('text',rep('numeric',cols-1),'skip')) %>%
            filter(label,startyear=startyear,endyear=endyear) %>%
            mutate(Year=as.numeric(yy)) %>% select(-yy) 
}
energy<-read_excel(bpfile,sheet=peccolname,skip=2,na='n/a',
        col_types=c('text',rep('numeric',colnrpec-1),'skip')) %>%
        energyFilter("Energy",startyear="1965",endyear=colpfxend)
```


```{r warning=FALSE}
#-------------------------------------------------------------------------------------------
# We will eventually need per cap Electricity by country
#-------------------------------------------------------------------------------------------
countryTWh<-bind_rows(
    getSheet(esheetname,colnr,'Electricity',startyear="1985"),
) %>% group_by(Country) %>% summarize(totalTWh=max(TWh)) 
countryTWh <- countryTWh %>% removeMissing(dfpop)
countryTotalMWhPerCap<-getPopForList(countryTWh$Country,2022) %>% left_join(countryTWh) %>%
  mutate(totalMWhPerCap=totalTWh*1e12/Popn/1e6) %>% # NB this is an annual figure
  select(Country,totalMWhPerCap)
# We pull out all the electricity data, and will decide what to keep later
# the data for electricity from Gas, Coal and Oil starts in 1985, instead of
# 1965, which is annoying
twh<-bind_rows(
    getSheet(gassheet,colnr,'Gas',startyear="1985"),
    getSheet(coalsheet,colnr,'Coal',startyear="1985"),
    getSheet(oilsheet,colnr,'Oil',startyear="1985"),
    getSheet("Solar Generation - TWh",colnrpec,'Solar'),
    getSheet("Wind Generation - TWh",colnrpec,'Wind'),
    getSheet("Hydro Generation - TWh",colnrpec,'Hydro'),
    getSheet("Nuclear Generation - TWh",colnrpec,'Nuclear'),
    getSheet("Geo Biomass Other - TWh",colnrpec,'Biomass/Geo')
  ) 
```


## Global Capacity Factors

```{r}
globalNuclearTWh<-twh %>% filter(Type=="Nuclear",Year==2023) %>% summarize(tworld=sum(TWh))
globalNuclearGW<-371.5   # From PRIS ... not in the statistical review

globalNuclearCapacityFactor<-(globalNuclearTWh*1e12)/(globalNuclearGW*1e9*24*365)

globalSolarTWh<-twh %>% filter(Type=="Solar",Year==2023) %>% summarize(tworld=sum(TWh))
globalSolarMW<-(solarWorld %>% filter(Year=="2023"))$MW

globalWindTWh<-twh %>% filter(Type=="Wind",Year==2023) %>% summarize(tworld=sum(TWh))
globalWindMW<-(windWorld %>% filter(Year=="2023"))$MW
globalWindCapacityFactor<-(globalWindTWh*1e12)/(globalWindMW*1e6*24*365)
globalWindCapacityFactor

globalSolarCapacityFactor<-(globalSolarTWh*1e12)/(globalSolarMW*1e6*24*365)
globalSolarCapacityFactor

print(paste0("Global capacity factors 2023: \n"))
print(paste0("       Wind: ",globalWindCapacityFactor,"\n"))
print(paste0("      Solar: ",globalSolarCapacityFactor,"\n"))
print(paste0("      Nuclear: ",globalNuclearCapacityFactor,"\n"))

solarToNuclear<-globalNuclearCapacityFactor/globalSolarCapacityFactor

solarTonnesPerGW=70000

print(paste0("GW of solar to generate nuclear TWh: ",globalNuclearGW*solarToNuclear,"\n"))
print(paste0("Tonnes of panels: ",globalNuclearGW*solarToNuclear*solarTonnesPerGW,"\n"))
```

## CF By Country

```{r}
solarTwhMW <- twh |> filter(Type=="Solar" & Year>=2000) 
solarCapLong <- solarCap |> pivot_longer(cols=`2000`:csyear,names_to="Year",values_to="MW") |> mutate(Year=as.numeric(Year))
cfSolar<-inner_join(solarTwhMW,solarCapLong,by=join_by(Country,Year)) |> mutate(CF=TWh/(MW*24*365/1e6))

windTwhMW <- twh |> filter(Type=="Wind" & Year>=2000) 
windCapLong <- windCap |> pivot_longer(cols=`2000`:csyear,names_to="Year",values_to="MW") |> mutate(Year=as.numeric(Year))
cfWind<-inner_join(windTwhMW,windCapLong,by=join_by(Country,Year)) |> mutate(CF=TWh/(MW*24*365/1e6))


#nuclearTwhMW <- twh |> filter(Type=="Nuclear" & Year>=2000) 
#nuclearCapLong <- nuclearCap |> pivot_longer(cols=`2000`:csyear,names_to="Year",values_to="MW") |> mutate(Year=as.numeric(Year))
#cfNuclear<-inner_join(nuclearTwhMW,nuclearCapLong,by=join_by(Country,Year)) |> mutate(CF=TWh/(MW*24*365/1e6))

```



## Plot per cap energy

```{r}
energyExj <- energy %>% mutate(Year=as.numeric(Year)) %>% removeMissing(dfpop)
energyByCountry<-energyExj %>% mergePopData(energyExj$Country %>% unique) %>% mutate(JPerCap=Exj*1e18/Popn)
energyByCountry %>% filter(Country %in% 
                             c("Canada","United Kingdom","China","Australia","Germany","US","France","Sweden")) %>%
                     ggplot() + geom_line(aes(x=Year,y=JPerCap/1e6,group=Country,color=Country)) +
            geom_smooth(aes(x=Year,y=JPerCap/1e6,group=Country,color=Country)) +
                    labs(title=paste0("Primary Energy per person (Data: Energy Institute WES ",theyearonly,")"),
                         x="",y="Megajoules per person")
write_csv(energyByCountry,paste0("energyByCountry",theyearonly,".csv"))

```

## Plotting wind, solar and hydro growth by country

```{r}
solarByCountry<-twh %>% filter(Type=="Solar",Year>=1995) %>% group_by(Country) %>% nest()
tmp <- solarByCountry %>% mutate(MaxTWh=map_dbl(data, ~ max(replace_na(.x$TWh,0),na.rm=TRUE))) %>% 
  arrange(desc(MaxTWh)) %>% ungroup() %>% slice(1:16) %>% unnest(data) 
```

```{r}
tmp <- tmp %>%  mergePopData(tmp$Country %>% unique) %>% 
  mutate(MWhPerCap=TWh*1e12/Popn/1e6) # NB this is over the period

p<-tmp %>% ggplot(aes(x=Year,y=MWhPerCap)) + geom_line() + facet_wrap(~Country) 
p
ttmp<-tmp %>% filter(!Country %in% c("Turkey","Vietnam")) %>%
  mutate(growth=100*(MWhPerCap-lag(MWhPerCap))/lag(MWhPerCap)) %>% filter(Year>=2000) 
p<-ttmp %>% ggplot(aes(x=Year,y=growth)) + 
  geom_line() + xlim(NA,as.numeric(theyearonly))+ labs(y="Growth Percentage",title="Solar growth rates in countries with the most solar power") +  
  geom_text(aes(x=Year,y=growth+100,label=comma(growth)),data=ttmp %>% filter(Year==2020)) + 
  facet_wrap(~Country) 
p
png(paste0("solar-growth-rates-by-top-countries-2000-2020-",theyear,".png"),width=1800,height=1800,units="px",res=200,type="cairo-png")
print(p)
dev.off()

```
## Plotting wind, solar and hydro growth globally

```{r}
#--------------------------------------------------------------------------
# cleaning up the data is complicated by missing values early in the series 
# so we start in 1998 and just plot from 2000
#--------------------------------------------------------------------------
calcAndClean<-function(.data) {
    mutate(.data,D1=TWh/lag(TWh)-1) %>%  
    mutate(d1=ifelse(is.infinite(D1)|is.nan(D1)|is.na(D1),0,D1)) %>% 
    mutate(percentGrowth=ifelse(d1>3,300,d1*100)) %>%
    mutate(percentGrowth=ifelse(abs(percentGrowth)>95,0,d1*100)) %>%
    select(-D1,-d1)  
}
# world totals without china
# xxxx<-getSheet("Solar Generation - TWh",60,'Solar',startyear="1998",filter=worldTotalFilterSubChina) %>%
#      mutate(dd=TWh-lag(TWh,23)) %>% filter(Country=="Total World") %>% 
#       select(Country,Type,Year,dd) %>% mutate(Country="WorldSubChina",TWh=dd) %>% select(-dd) 
```

```{r}
selectedClist=c("Total Europe","Germany","Denmark","Sweden","Australia","Spain")
selectedLabel="EuDeuDenSweAusSpa"
makeFilter<-function(cl) {
       function(.data,name,startyear="1975",CL=cl,endyear=cfilend) {
        rename(.data,Country=`Terawatt-hours`) %>% mutate(Type=name) %>% filter(!is.na(Country)) %>% 
        filter(Country %in% CL) %>% slice(1:84) %>% 
        select(Country,'Type',startyear:endyear) %>% rename(!!csyear:=endyear) %>% 
        pivot_longer(startyear:!!csyear,names_to="yy",values_to="TWh") %>% filter(!is.na(TWh))
       }
}
selectedFilter<-makeFilter(selectedClist)

stwselected<-bind_rows(
    getSheet("Solar Generation - TWh",colnrpec,'Solar',startyear="1998",filter=selectedFilter),
    getSheet("Wind Generation - TWh",colnrpec,'Wind',startyear="1998",filter=selectedFilter),
    getSheet("Hydro Generation - TWh",colnrpec,'Hydro',startyear="1998",filter=selectedFilter)
) %>% calcAndClean()
stwsubch<-bind_rows(
    getSheet("Solar Generation - TWh",colnrpec,'Solar',startyear="1998",filter=worldTotalFilterSubChina) %>% 
      mutate(dd=TWh-lag(TWh,23)) %>% filter(Country=="Total World") %>%
      select(Country,Type,Year,dd) %>% mutate(Country="WorldSubChina",TWh=dd) %>% select(-dd),
    getSheet("Wind Generation - TWh",colnrpec,'Wind',startyear="1998",filter=worldTotalFilter),
    getSheet("Hydro Generation - TWh",colnrpec,'Hydro',startyear="1998",filter=worldTotalFilter)
) %>% calcAndClean()
# world totals 
stwh<-bind_rows(
    getSheet("Solar Generation - TWh",colnrpec,'Solar',startyear="1998",filter=worldTotalFilter),
    getSheet("Wind Generation - TWh",colnrpec,'Wind',startyear="1998",filter=worldTotalFilter),
    getSheet("Hydro Generation - TWh",colnrpec,'Hydro',startyear="1998",filter=worldTotalFilter)
) %>% calcAndClean()
calcGrowth<-function(.data) {
  .data %>% ggplot() + geom_line(aes(x=Year,y=TWh,color=Type,group=Type),linewidth=2)+
  labs(title="Growth of hydro, solar and wind",y="Terawatt-hours")+
  annotate('text',x=2000,y=300,hjust=0,label=paste0("Data: Energy Institute ",theyearonly),size=3)+
  theme(axis.title=element_text(size=20,face="bold"),
        axis.text.x=element_text(size=20,face="bold"),
        plot.title=element_text(size=20,face="bold"),
        axis.text.y=element_text(size=20,face="bold"),
        legend.title=element_text(size=20,face="bold"),
        legend.text=element_text(size=20,face="bold")
        )+xlim(2000,as.numeric(theyearonly)-1)
}
calcGrowthRatePanel<-function(.data) {
  .data %>% ggplot() + 
  geom_line(aes(x=Year,y=percentGrowth,color=Type,group=Type),linewidth=0.5)+
    facet_wrap(Country~.)+
  labs(title=paste0("Percentage growth rates of hydro, solar and wind (Data:Energy Institute ",theyearonly,")"),y="Percentage growth")+
  theme(axis.title=element_text(size=4,face="bold"),
        axis.text.x=element_text(size=4),
        plot.title=element_text(size=4),
        axis.text.y=element_text(size=4),
        legend.title=element_text(size=4),
        legend.text=element_text(size=4)
        )+
  xlim(c(2000,as.numeric(theyearonly)-1))
}
calcGrowthRate<-function(.data,suffix) {
  .data %>% ggplot() + 
  geom_line(aes(x=Year,y=percentGrowth,color=Type,group=Type),linewidth=2)+
  labs(title=paste("Percentage growth rates of hydro, solar and wind",suffix),y="Percentage growth")+
  annotate('text',x=2005,y=10,label=paste0("Data: Energy Institute ",theyearonly),size=3)+
  theme(axis.title=element_text(size=12,face="bold"),
        axis.text.x=element_text(size=12,face="bold"),
        plot.title=element_text(size=12,face="bold"),
        axis.text.y=element_text(size=12,face="bold"),
        legend.title=element_text(size=12,face="bold"),
        legend.text=element_text(size=12,face="bold")
        )+
  xlim(c(2000,as.numeric(theyearonly)-1))
}
wshGrowth<-calcGrowth(stwh)

sgr<-calcGrowthRatePanel(stwselected)
sgr
png(paste0("selectedCompoundGrowthRates-",selectedLabel,"-",theyear,".png"),width=1200,height=1200,units="px",res=300,type="cairo-png")
sgr
dev.off()

wshGrowthRate<-calcGrowthRate(stwh,"(global)")

wshGrowthRate

# and without china
wgrsubc<-calcGrowthRate(stwsubch,"(without China)")
wgrsubc

png(paste0("wsh-growth-2000-2020-",theyear,".png"),width=1800,height=1300,units="px",res=200,type="cairo-png")
print(wshGrowth)
dev.off()
png(paste0("wsh-growthrate-2000-2020-",theyear,".png"),width=2600,height=1300,units="px",res=200,type="cairo-png")
print(wshGrowthRate)
dev.off()

```

```{r}
#-----------------------------------------------
# Augment twh with sums of various columns
#-----------------------------------------------
```
```{r}
twhEx<-twh %>% pivot_wider(names_from=c("Type"),values_from="TWh") %>% replace(is.na(.),0) %>%
  mutate(Clean=Solar+Wind+Nuclear+Hydro,SolarPlusWind=Solar+Wind,SolarPlusWindPlusBio=Solar+Wind+`Biomass/Geo`) %>%
  pivot_longer(cols=`Gas`:`SolarPlusWindPlusBio`,names_to="Type",values_to="TWh")
```

```{r}
#-----------------------------------------------
# Remove Countries for which we have no population figures 
# and build a country list
#-----------------------------------------------
twh <- twh %>% removeMissing(dfpop) 
clist<-twh %>% select(Country) %>% unique()
#---------------------------------------------------------------------------
# Now we can augment the data with population and calculate perCap numbers
#---------------------------------------------------------------------------
fulldfEx<-twhEx %>% mergePopData(clist$Country) %>%  filter(LocTypeName=="Country/Area") %>%
  mutate(MWhPerCap=TWh*1e12/Popn/1e6) %>% arrange(Country,Year,Type) %>% select(-LocTypeName) # NB this is over the period
```

```{r}
#-------------------------------------------------------------------------
# Now we nest the data for each Country,Type
#-------------------------------------------------------------------------
nst<-fulldfEx %>% group_by(Country,Type) %>% arrange(Year) %>% nest()
#-------------------------------------------------------------------------
# Find the clean energy per cap of year
#-------------------------------------------------------------------------
cleanData <- fulldfEx %>% filter(Type=="Clean",Year==as.numeric(theyearonly)-1) %>% 
      select(Country,TWh,Popn) %>% mutate(cleanMWhPerCap=TWh*1e12/Popn/1e6) %>% select(-TWh,-Popn)
```

```{r}
pickCountries<-function(n) {
  dfpop %>% filter(Year==theyearonly,Popn>n) %>% select(Country)
}
cc<-pickCountries(26e6)
ccc<-nst %>% semi_join(cc) %>% filter(!is.na(Type))

```

```{r}
#----------------------------------------------------------
# Now we can calculate the lag increment for each country and type
# how much per capita energy they added over a specified period
# NB slice only works on ungrouped data 
#----------------------------------------------------------
#xxx<-data_frame()
#yyy<-data_frame()
makeRollingGraph<-function(rp,ntop=30,rmTypes=c("SolarPlusWindPlusBio","Clean"),poplimit=0) {
    fnst=nst
    if (poplimit>0) {
       cc<-pickCountries(poplimit)
       ccc<-fnst %>% semi_join(cc) %>% filter(!is.na(Type))
       fnst=ccc
    }
    mnst<-fnst %>% mutate(data=map(data,~mutate(.x,inc=.$MWhPerCap-lag(.$MWhPerCap,rp)))) 
    mnstmax<-mnst %>% mutate(Max=map_dbl(data, ~ max(replace_na(.x$inc,0),na.rm=TRUE)),
                         MaxYear=map(data, ~ .$Year[which.max(replace_na(.x$inc,0))]),
                         MaxPop=map(data, ~ ceiling(.$Popn[which.max(replace_na(.x$inc,0))]))
                         ) %>% left_join(countryTotalMWhPerCap)
    mnstmax<-mnstmax %>% left_join(cleanData)
    #xxx<<-mnstmax
    plim=""
    if (poplimit>0) {
        plim=paste0("over",poplimit/1e6,"m-")
    }
    write_csv(mnstmax %>% unnest(c(data,MaxYear,MaxPop)),file=paste0("growthtables",plim,theyear,"-",rp,".csv"))
    saveRDS(mnstmax,file=paste0("growthtables",theyear,"-",rp,".rds"))
    # dbg: write_csv(nst %>% unnest() %>% filter(Country=="Iceland"),paste0("unnested",rp,"-",ntop,".csv"))
    # dbg: write_csv(mnstmax %>% unnest() %>% filter(Country=="Iceland"),paste0("unnestedmax",rp,"-",ntop,".csv"))
    topn<-mnstmax %>% filter(!Type %in% rmTypes) %>% 
      arrange(desc(Max)) %>% 
      select(Country,Type,MaxYear,Max,MaxPop,cleanMWhPerCap,totalMWhPerCap) %>% ungroup() %>% 
        slice_head(n=ntop) %>% 
        mutate(Rank=rank(desc(Max))) 
    colours<-c("Hydro"="#3399FF","Nuclear"="purple","Solar"="yellow","SolarPlusWind"="orange","Wind"="coral1","Coal"="black","Gas"="grey")
    #yyy<<-topn
    p<-topn %>% mutate(Key=paste0(Rank,". ",Country,", ",MaxYear,", ",comma2(as.numeric(MaxPop)/1e6),"m")) %>%
      mutate(therow=reorder(Key,Max)) %>%
      ggplot(aes(x=therow,y=Max))+
      geom_col(alpha=1,aes(fill=Type))+
      geom_point(aes(y=totalMWhPerCap*2,shape="2050 Target"),alpha=0.8,size=2) +
      geom_point(aes(y=totalMWhPerCap,shape="Current MWh/Cap"),alpha=0.8,size=1) +
      geom_point(aes(y=cleanMWhPerCap,shape="Current Clean"),alpha=0.8,size=1) +
      geom_text(aes(y=Max+1,label=paste0(comma2(Max)),hjust=0),size=3)+ 
      coord_flip()+
      scale_fill_manual(values=colours)+
      labs(x="",title=paste0("Highest megawatt-hours per person added in ",rp," years"),
           y="MWh per person over period")+ylim(c(0,40))+
      annotate('text',x=2,y=20,hjust=0,label=paste0("Data: Energy Institute ",theyearonly),size=2)+
      scale_shape_manual(values=c(19,8,1),labels=c("2050 Target", "Current Clean", "Current MWh/Cap"
                                                   ))+
      guides(shape=guide_legend("Targets"))
      suf=""
      if (length(rmTypes)>2) {
        suf=""
        p<-p+ geom_rect(aes(ymin=cleanMWhPerCap,ymax=totalMWhPerCap*2,
              xmax=as.numeric(factor(reorder(Key,Max)))+0.4,
              xmin=as.numeric(factor(reorder(Key,Max)))-0.4),alpha=0.3,fill="forestgreen") +
            geom_text(aes(y=totalMWhPerCap*2+1,
            label=paste0(comma2((totalMWhPerCap*2-cleanMWhPerCap)/Max),"x",rp," yrs"),hjust=0),size=3)
      }
    
      t<-fulldfEx %>% select(Type) %>% unique() %>% filter(!Type %in% rmTypes) %>% map(~substr(.,1,1))
      t<-paste(unlist(t),sep="",collapse="")
      plim=""
      if (poplimit>0) {
        plim=paste0("over",poplimit/1e6,"m-")
      }
      png(paste0("rollout-top-",plim,theyear,"-",ntop,"-over-",rp,"-years-",t,".png"),width=2200,height=1300,units="px",res=300,type="cairo-png")
      print(p)
      dev.off()
      print(p)
}
makeRollingGraph(7,ntop=20)
makeRollingGraph(10,ntop=30,rmTypes=c("Gas","Coal","SolarPlusWindPlusBio","Clean","Oil"))
makeRollingGraph(10,ntop=30,rmTypes=c("Gas","Coal","SolarPlusWindPlusBio","Clean","Oil"),poplimit=20e6)
makeRollingGraph(10,ntop=30,poplimit=20e6)
```
```{r}
makeRollingGraph(15,ntop=30)
makeRollingGraph(20,ntop=20)
makeRollingGraph(20,ntop=20,rmTypes=c("Gas","Coal","SolarPlusWindPlusBio","Clean","Oil"))
makeRollingGraph(10,ntop=30)
makeRollingGraph(10,ntop=30,rmTypes=c("Gas","Coal","SolarPlusWindPlusBio","Clean","Oil"))
makeRollingGraph(10,ntop=30,rmTypes=c("Gas","Coal","SolarPlusWindPlusBio","Clean","Oil"),poplimit=20e6)
```


```{r}
#-------------------------------------------------------------------------
# At this point we could go searching for top countries, but for now we just
# select interesting ones
#-------------------------------------------------------------------------
mnst<-nst %>% mutate(data=map(data,~mutate(.x,inc=.$MWhPerCap-lag(.$MWhPerCap,5)))) 
selectedCountries<-mnst %>% unnest(cols=c(data)) %>% 
  filter(Country %in% c("France","Germany","US","Sweden","Belgium","United Kingdom")) %>%
  filter(Type %in% c("Nuclear","SolarPlusWind","SolarPlusWindPlusBio","Hydro")) 
p<-selectedCountries %>% ggplot(aes(x=Year,y=inc,color=Type)) + geom_line(linewidth=0.3) + facet_wrap(~Country) + 
  labs(y="Increase in Megawatt-hours per person over 5 years",title="Electrical energy change by source (data: BP Stats 2021)")+
  theme(axis.title=element_text(size=7),axis.text.x=element_text(size=7))
p
png(paste0("megawatt-hours-per-person-increase-",theyear,".png"),width=1800,height=800,units="px",res=300,type="cairo-png")
p
dev.off()
```

```{r}
#---------------------------------------------------------
# The above graphs are better, but this next is interesting
# because of it's use of two legends in the graph. It also
# has the advantage of not showing things relative to zero;
# which can confuse some people.
#---------------------------------------------------------
# twh has data for all countries and mergePopData will filter out
# a selected set and add in the population data
#---------------------------------------------------------
ctwh<-twh %>% mergePopData(c("Australia","China","US","France","Japan","Germany","Belgium","Sweden","United Kingdom"))
totaltwh <- ctwh %>% group_by(Country,Year) %>% summarise(total=sum(TWh)) %>% 
  select(Country,Year,total) %>% left_join(ctwh %>% select(Country,Year,Popn) %>% unique()) 

p<-ctwh %>% filter(!Type %in% c("Gas","Coal","Oil")) %>% ggplot(aes(x=Year,y=TWh*1e12/Popn/1e6))+
  geom_col(aes(fill=Type))+geom_line(aes(x=Year,y=total*1e12/Popn/1e6,color="black"),data=totaltwh)+
  guides(fill=guide_legend("Show only"),color=guide_legend(""))+
  scale_color_manual(values=c("black"),labels="Total")+
  xlim(1985,as.numeric(theyearonly)-1)+
  facet_wrap(~Country)+
  labs(y="Megawatt-hours per capita from clean(ish) energy sources",
       title=paste0("Clean vs total electricity (Data: Energy Institute Stats (",theyearonly,")"))+
  theme(axis.title=element_text(size=7),axis.text.x=element_text(size=7))
p
#--------------------------------------------------------------------------------------
# now make a png for posting
#--------------------------------------------------------------------------------------
png(paste0("cleanish-megawatt-hours-per-person-",theyear,".png"),width=1800,height=1200,units="px",res=300,type="cairo-png")
p
dev.off()
```


