---
title: "Redo Thorcon chart"
author: "GeoffRussell"
date: "29/10/2021"
output:
  pdf_document: default
  html_document: default
---

## Background

Thorcon have a great chart, first version by Robert Hargraves and redone by Jack Devanney, which shows
the various electricity inequalities between different regions.  This *markdown* file will
do yet another version using BP World Energy Statistics 2021. Most of the code here comes from
other analyses I've done of the BP data. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(readxl)
library(plotly)
library(ggrepel)
library(RColorBrewer)
comma<-function(x) prettyNum(signif(x,digits=4),big.mark=",")
comma3<-function(x) prettyNum(signif(x,digits=3),big.mark=",")
comma2<-function(x) prettyNum(signif(x,digits=2),big.mark=",")
```

First read the data. Hopefully, we can just run this script again when the next version of the BP statistics
are released.  We use the 2020 global population figure from [Worldometers](https://www.worldometers.info/world-population/world-population-by-year/).


```{r data}
bpfile<-"bp-stats-review-2021-all-data.xlsx"
s21<-excel_sheets(bpfile)
print(s21)
worldPop20<-7.795e9
```

## Population data

We will be calculating per-capita statistics so need some population data.

We use a download of the 2020 [World Bank Population data](https://datahelpdesk.worldbank.org/knowledgebase/articles/898581-api-basic-call-structures) 
which gives population time series data upto 2019. We use a polynomial model
to extrapolate for years after that; out to 2040. It's not the world's best population growth model, but good
enough for our purposes.

One wrinkle is that the World Bank country names don't match the BP country names. So we change
some names. But the BP data is still messy; using categories like *Eastern Africa*, for example,
which combine a group of countries.

```{r pop}
dfpop<-read_csv("API_SP.POP.TOTL_DS2_en_csv_v2_2252106.csv",skip=4,show_col_types=FALSE) %>% rename(Country=`Country Name`)
dfpop$Country <- gsub("United States","US",dfpop$Country)
dfpop$Country <- gsub("Korea, Rep.","South Korea",dfpop$Country)
dfpop$Country <- gsub("Slovak Republic","Slovakia",dfpop$Country)
dfpop$Country <- gsub("Egypt, Arab Rep.","Egypt",dfpop$Country)
dfpop$Country <- gsub("Hong Kong SAR, China","China Hong Kong SAR",dfpop$Country)
dfpop$Country <- gsub("Iran, Islamic Rep.","Iran",dfpop$Country)
dfpop$Country <- gsub("Venezuela, RB","Venezuela",dfpop$Country)
dfpop$Country <- gsub("Trinidad and Tobago","Trinidad & Tobago",dfpop$Country)
```


```{r}
#------------------------------------------------------------------------------------------------
# getpop returns a closure which is vectorised, we extrapolate past the last data point, 2019
# we unname pred because we don't need names and they confuse me when looking at structures!
#------------------------------------------------------------------------------------------------
#----------------------------------------------------------------------
# Remove any countries in twh for which we don't have population data
#---------------------------------------------------------------------
removeMissing<-function(.data,df) {
  semi_join(.data,df %>% select(Country),by=c("Country")) 
}
getpop<-function(country) {
  f<-function(yr) {
    rp<-dfpop %>% filter(Country==country) %>% pivot_longer(names_to="year",cols="1960":"2019") %>% select(year,value)
    m1<-lm(value~poly(as.numeric(year),2),rp)
    df<-tibble(year=seq(1960,2040),Country=country)
    rowp<-df %>% add_predictions(m1) %>% filter(year %in% yr) %>% mutate(Year=year,Popn=unname(pred)) %>% select(-year,-pred) 
  }
}
getPopForList<-function(countries,years=seq(1975,2021)) {
  df<-tibble()
  for(c in countries) {
    df<-bind_rows(df,getpop(c)(years))
  }
  df$Popn<-unname(df$Popn)
  df 
}
getTotalTWhForList<-function(countries,years=seq(1975,2021)) {
  df<-tibble()
  for(c in countries) {
    df<-bind_rows(df,(countryTWh %>% filter(Country==c)))
  }
  df$TWh<-unname(df$TWh)
  df 
}
# Here's a couple of test calls
tgetpop<-getpop("Australia")(1975:2000)
tgetPopForList<-getPopForList(c("Australia","Germany","France","US"))
mergePopData<-function(.data,countries,years=c(2021)) {
    dfc<-getPopForList(countries,years)
    .data %>% filter(Country %in% countries) %>% left_join(dfc)
}
mergeTotalTWhData<-function(.data,countries) {
    dfe<-getTotalTWhForList(countries)
    .data %>% filter(Country %in% countries) %>% left_join(dfe)
}
```
## Get data for a sheet and filter it

We now have a set of routines for reading a BP sheet and filtering it. This is all a bit
messy because of the way BP name things.

```{r }
# plenty of data juggling due to how read_excel names its columns
# this code will have to change in 2022!
countryFilter<-function(.data,name,startyear="1975",endyear="2020...57") {
  rename(.data,Country=`Terawatt-hours`) %>% mutate(Type=name) %>% filter(!is.na(Country)) %>% 
    filter(!grepl("^Total ",Country)) %>%  
    filter(!grepl("OECD",Country)) %>%  
    filter(!grepl("European Union",Country)) %>%  
    filter(!grepl("^of which",Country)) %>% 
    select(Country,'Type',startyear:endyear) %>% rename('2020'=endyear) %>% 
    pivot_longer(startyear:'2020',names_to="yy",values_to="TWh") %>% filter(!is.na(TWh))
}
worldTotalFilter<-function(.data,name,startyear="1975",endyear="2020...57") {
  rename(.data,Country=`Terawatt-hours`) %>% mutate(Type=name) %>% filter(!is.na(Country)) %>% 
    filter(grepl("^Total World",Country))  %>% 
    select(Country,'Type',startyear:endyear) %>% rename('2020'=endyear) %>% 
    pivot_longer(startyear:'2020',names_to="yy",values_to="TWh") %>% filter(!is.na(TWh))
}
getSheet<-function(sheetname,cols,label,startyear="1975",filter=countryFilter) {
    endyear<-paste0("2020...",toString(cols-3))
    read_excel(bpfile,sheet=sheetname,skip=2,na='n/a',col_types=c('text',rep('numeric',cols-1))) %>%
            filter(label,startyear=startyear,endyear=endyear) %>%
            mutate(Year=as.numeric(yy)) %>% select(-yy) 
}
```

We won't to get per capita electricity by country, but the BP data has plenty of data rows with
country names like *Other SE Asia*, *Other South Africa*. These block up 
countries with little electricity; e.g., most of Africa. So we will group the data into *Country*
data and *Other* data. 

We will total up the population covered by the *Country* data and then just allocate the rest of
the global population to the *Other* data.

```{r}
countryTWh<-bind_rows(
    getSheet("Electricity Generation ",40,'Electricity',startyear="1985"),
) %>% filter(Year==2020) %>% select(Country,TWh)
write_csv(countryTWh,"country-data.csv")
countryTWh %>% summarise(s=sum(TWh))
otherTWh <- countryTWh %>% filter(grepl("Other|Eastern Africa|Middle Africa|Central America|Western Africa",Country))
write_csv(otherTWh,"other-data.csv")
otherSum<-otherTWh %>% summarise(sum(TWh))

# at this point countryTWh still has Other data ... so we remove all countries for which
# we don't have data; which will hopefully only be the Other data; but we write out a file to check.

countryTWh <- countryTWh %>% removeMissing(dfpop)
write_csv(countryTWh,"country-data-stripped.csv")
countryTotalMWhPerCap<-getPopForList(countryTWh$Country,2021) %>% left_join(countryTWh) %>%
  mutate(totalMWhPerCap=TWh*1e12/Popn/1e6) %>% # NB this is an annual figure
  select(Country,totalMWhPerCap)
```

The two csv files are just so we can check by eye that nothing has been missed.

```{r}
countryTWh %>% summarise(s=sum(TWh))
```
The `countryTWh` and `otherTWh` have the country based data and the `Other` data. 

Our basic method is to place the countries in categories and allocate the `Other` data to
a *RestOfWorld* bucket.

## Group countries in regions

Now we read in a table of regions which has been manually doctored to allocate countries to regions.
The manual allocation tends to put like with like; based on electrical availability. So Australia
is allocated to *North America* and Mexico to *South America*. The richer middle eastern countries
are allocated to *Europe*.


```{r}
regions<-read_csv("regionTable.csv") %>% select(Country,Region)
```

Add in the population and region data.

```{r}
c<-countryTWh$Country
noregion<-countryTWh %>% anti_join(regions %>% select(Country),by=c("Country"))
dfdata<-countryTWh %>% mergePopData(c) %>% left_join(regions) %>% 
  mutate(PerCapEnergy=TWh*1e12/Popn/24/365)
dfdata %>% summarise(wPop=sum(Popn))
dfg<-dfdata %>% group_by(Region) %>%
  summarise(Population=sum(Popn)/1e6,PerCapEnergy=sum(TWh*1e12)/sum(Popn)/24/365) %>%
  arrange(desc(PerCapEnergy))
dfg %>% summarise(wPop=sum(Population)) 
```

At this point `dfg` is missing the *other* information.

```{r}
missingPop<-(worldPop20/1e6-as.numeric(dfg %>% summarise(sum(Population))))*1e6
missingTWh<-as.numeric(otherSum)
restOfWorld<-tribble(
    ~Region,~Population,~PerCapEnergy, 
    "RestOfWorld",missingPop/1e6,missingTWh*1e12/missingPop/24/365
)
```

```{r}
dfg<-bind_rows(dfg,restOfWorld)
etarget<-780
bigRegionList<-c("North America",
              "Europe",
              "China",
              "Asia",
              "SE Asia",
              "South America"
              )
regionList<-c("North America",
              "Europe",
              "China",
              "Asia",
              "South America",
              "SE Asia",
              "India",
              "RestOfWorld"
              )
df <- dfg %>% mutate(xmin=cumsum(Population)-Population,xmax=xmin+Population,ymin=0,ymax=PerCapEnergy) 
dfepoor <- df %>% filter(!Region %in% c("North America","Europe")) 
shortfall<-dfepoor %>% summarise(sum((etarget-PerCapEnergy)*Population/1000))
totalgw<-df %>% summarise(sum(PerCapEnergy*Population/1000))
dfbig <- df %>% filter(Region %in% bigRegionList)
                                     
dfsmall <- df %>% filter(!Region %in% bigRegionList)
                           
dfr<-tribble(
  ~xmin,~xmax,~ymin,~ymax,
   0,    worldPop20/1e6,0,etarget
)
p<-df %>% ggplot()+
  geom_rect(aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),fill="grey85",data=dfr) +
  geom_rect(aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax,fill=fct_reorder(Region,PerCapEnergy))) +
  geom_text(aes(x=xmin+(xmax-xmin)/2,y=ymin+(ymax-ymin)/2,
                label=paste0(comma(PerCapEnergy*Population/1000),"GW")),data=dfbig,angle=90,size=2.3,color="white") +
  geom_text(aes(x=xmin+(xmax-xmin)/2,y=ymin+(ymax-ymin)/2,
                label=paste0(comma(PerCapEnergy*Population/1000),"GW")),data=dfsmall,size=3) +
  geom_text(aes(x=xmin+60,y=ymax+110,label=Region,hjust=0),size=2)+
  geom_text(aes(x=xmin+60,y=ymax+70,label=paste0(comma(PerCapEnergy),"w"),hjust=0),size=2)+
  geom_text(aes(x=xmin+60,y=ymax+30,label=paste0(comma(Population),"m"),hjust=0),size=2)+
  annotate('text',x=1000,y=1500,label=paste0("Current Global Power: ",comma(totalgw),"GW"),hjust=0,vjust=0)+
  annotate('text',x=100,y=20,label=paste0("Data: BP World Energy Statistics 2021 (World Bank Population data)"),hjust=0,vjust=0,size=2)+
  annotate('text',x=7500,y=700,label=paste0("Developing Country Shortfall: ",comma(shortfall),"GW"),hjust=1)+
  scale_fill_brewer(breaks=regionList,type="seq",palette="OrRd")+
  labs(y="Watts (continuous power) per person",x="Population",fill="Region",title="Global Electricity Shortfall (cf. Thorcon website)")
p
png(paste0("thorcon-pop-energy-new.png"),width=1800,height=900,units="px",res=200,type="cairo-png")
p
dev.off()
```
We are eventually going to need to list all the countries in each region.
```{r}
# this breaks of lists into lines of about 40 chars at a blank
splitter<- function(s) {
    gsub('(.{1,40})(\\s|$)', '\\1\n', s)
}
regionCountries<-regions %>% group_by(Region) %>% 
  mutate(str=splitter(paste(Country,sep=", ",collapse=", "))) %>% 
  ungroup() %>% select(Region,str) %>% unique()
```

```{r}
# This builds a bunch of rectangle,
# we don't need it because plotly allows bars of different widths, but its a very 
# useful trick for looping through rows
l<-list()
l<-pdf %>% pmap(
  function(...) {
    c<-tibble(...)
    l<-append(l,list(type="rect",text=c$Region,hoverinfo="text",x0=c$xmin,x1=c$xmax,
                     y0=c$ymin,y1=c$ymax,fillcolor=c$colr,opacity=0.2,line=list(width=0)))
  }
)
```

Now lets try a plotly version.

```{r}
pdf<-df %>% left_join(regionCountries) %>% mutate(colr=rev(getPalette(nrow(df)))) %>% 
  mutate(x=xmin+(xmax-xmin)/2,y=ymax,width=xmax-xmin)
fig<-plot_ly()
fb <- list(size = 13, color = "black")
fw <- list(size = 18, color = "grey")
fbig <- list(size = 18)
etarget=1200
popm<-worldPop20/1e6
pop50<-9600
fig<-layout(fig,title="TITLE",showlegend=FALSE,
            shapes=list(x0=0,y0=0,x1=popm,y1=etarget,layer="below",
                        fillcolor="Khaki",type="rect",line=list(width=0)),
            xaxis=list(title="Global Population (millions)"),
            yaxis=list(title="Watts per perons (continuous power)")) %>%
            layout(annotations=list(text=paste0("Target average global continuous\npower ",
                                etarget," watts per person"),x=2000,y=etarget,font=fb)) %>%
            layout(annotations=list(text=paste0("Global power required in 2050 with\n",
                                                comma(pop50/1000)," billion people: ",
                            comma(etarget*pop50/1000)," GW"),x=5500,y=etarget-110,showarrow=FALSE,font=fw))
            
fig<-fig %>% add_trace(data=pdf,type="bar",x=~x,y=~y,width=~width,text=~Region,hovertext=~str,
                       textfont=fbig,
                       hovertemplate="W/Cap: %{y}\n%{hovertext}<extra></extra>",
                       marker=list(color=~colr))
fig
htmlwidgets::saveWidget(as_widget(fig), "thorconplot.html")

```

Possibly a better way of dividing countries is simply ranking by electrical energy per person. Perhaps
splitting the world into population quintiles will work. That will enable us to compare the
electricity used by the richest (in electricity usage) 20 percent of the global population with
the rest.



```{r}
dfdata<-dfdata %>% arrange(-PerCapEnergy)
top15<-dfdata %>% slice(1:15)
ndiv=7
dgroups<-dfdata %>% mutate(Quantile=paste0("G",ntile(-PerCapEnergy,ndiv)))
groups<-dgroups %>% select(Country,Quantile,PerCapEnergy)
qdfg<-dgroups %>% group_by(Quantile) %>%
  summarise(Population=sum(Popn)/1e6,PerCapEnergy=sum(TWh*1e12)/sum(Popn)/24/365) %>%
  arrange(desc(PerCapEnergy)) %>% rename(Region=Quantile)
qdfg<-bind_rows(qdfg,restOfWorld)
dfgroups <- qdfg %>% mutate(xmin=cumsum(Population)-Population,xmax=xmin+Population,ymin=0,ymax=PerCapEnergy) 
qList<-unlist(d$Quantile %>% unique())
qList<-append(qList,"RestOfWorld")
getPalette = colorRampPalette(brewer.pal(9, "OrRd"))

totalgw<-dfgroups %>% summarise(sum(PerCapEnergy*Population/1000))
totalpop<-dfgroups %>% summarise(sum(Population))
dfbig <- dfgroups %>% filter(Region %in% c("G1","G2","G3"))
 
etarget<-dfgroups$PerCapEnergy[1]/2
dfr<-tribble(
  ~xmin,~xmax,~ymin,~ymax,
   0,    worldPop20/1e6,0,etarget
)
top15<-top15 %>% mutate(y=1550-seq(1:15)*50)
dfepoor <- dfgroups %>% filter(!Region %in% c("G1","G2"))
shortfall<-dfepoor %>% summarise(sum((etarget-PerCapEnergy)*Population/1000))
p<-dfgroups %>% ggplot()+
  geom_rect(aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),fill="grey85",data=dfr) +
  annotate('text',x=1000,y=1400,label=paste0("Current Global Power: ",comma(totalgw),"GW"),hjust=0,vjust=1)+
  geom_rect(aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax,fill=fct_reorder(Region,PerCapEnergy))) +
  annotate('text',x=7500,y=1600,label=paste0("Top 15"),hjust=1)+
  geom_text(aes(x=7500,y=y,
                label=paste0(Country," ",comma(floor(PerCapEnergy)),"w")),
            hjust=1,size=3,data=top15)+
  geom_text(aes(x=xmin+(xmax-xmin)/2,y=ymin+(ymax-ymin)/2,
                label=paste0(comma(floor(PerCapEnergy*Population/1000)),
                             "GW (",comma(ceiling(100*Population/as.numeric(totalpop))),"%Pop)")),
            data=dfbig,angle=90,size=3,color="white") +
  annotate('text',x=7500,y=700,label=paste0("Developing Country Shortfall: ",comma(shortfall),"GW"),hjust=1)+
  geom_text(aes(x=xmin+60,y=ymax+30,label=paste0(comma(Population),"m"),hjust=0),size=2)+
  geom_text(aes(x=xmin+60,y=ymax+70,label=paste0(comma(PerCapEnergy),"w"),hjust=0),size=2)+
  scale_fill_manual(breaks=qList,values=rev(getPalette(ndiv+1)))+
  labs(y="Watts (continuous power) per person",
       x=paste0("Population (",comma(as.numeric(totalpop)),"m)"),
       fill="Region",title="Global Electricity Shortfall (cf. Thorcon website)")
  
p
png(paste0("thorcon-pop-energy-new-octile.png"),width=1800,height=900,units="px",res=200,type="cairo-png")
p
dev.off()
```

This seems, dividing the world into 8 levels of electricity use, to show the inequalities
most clearly. 

Now let's try that in plotly.

```{r}
groupCountries<-groups %>% group_by(Quantile) %>% 
  mutate(str=splitter(paste(paste0(Country," ",comma(PerCapEnergy),"w"),sep=", ",collapse=", "))) %>% 
  ungroup() %>% select(Quantile,str) %>% rename(Region=Quantile) %>% unique() 
pdfgroups<-dfgroups %>% left_join(groupCountries) %>% mutate(colr=rev(getPalette(nrow(df)))) %>% 
  mutate(x=xmin+(xmax-xmin)/2,y=ymax,width=xmax-xmin)
pdfgroups$str<-ifelse(is.na(pdfgroups$str),"Many poor countries have no\nindividual data in\nthe BP World Energy Statistics\nE.g. Nigeria with over 200 million people",pdfgroups$str)
pdfgroups<-pdfgroups %>% mutate(str=paste0("Popn: ",comma(Population),"m\n",str))
gfig<-plot_ly()
fb <- list(size = 13, color = "black")
fw <- list(size = 18, color = "grey")
fbig <- list(size = 18)
etarget=1200
popm<-worldPop20/1e6
pop50<-9600
gfig<-layout(gfig,title=list(
          text="<b>Global per person distribution of electricity 2020</b>",
          y=0.90,x=0.2, xanchor='left',yanchor='top',font=list(size=20)
          ),showlegend=FALSE,
            shapes=list(x0=0,y0=0,x1=popm,y1=etarget,layer="below",
                        fillcolor="Khaki",type="rect",line=list(width=0)),
            xaxis=list(title="Global Population (millions)"),
            yaxis=list(title="Watts per person (continuous power)")) %>%
            layout(margin=list(t=80)) %>%
            layout(annotations=list(
              text=paste0("Possible scenario... ",etarget," watts per person"),
              hovertext=paste0("With target average global continuous\npower ",
                                etarget," watts per person ...\n",
          "global power required in 2050 with\n",
              comma(pop50/1000)," billion people: ",
              comma(etarget*pop50/1000)," GW"
                                ),align='left',showarrow=TRUE,x=2200,y=etarget,font=fb)) %>%
            layout(annotations=list(
              text=paste0("Data: BP World Energy Stats (Population data from World Bank)"),xanchor='left',x=100,y=60,font=list(color='white'),showarrow=FALSE)) %>%
            layout(annotations=list(hovertext="IEA 2050 target is 2.6 times\nthis present electricity output",
              text=paste0("Current global power with\n",
              comma(popm/1000)," billion people: ",comma(totalgw)," GW"),font=list(size="12"),
              x=5500,y=300,xanchor='left',showarrow=FALSE))
            
gfig<-gfig %>% add_trace(data=pdfgroups,type="bar",x=~x,y=~y,width=~width,text=~Region,hovertext=~str,
                       textfont=fbig,
                       hovertemplate="Avg w/Cap: %{y}\n%{hovertext}<extra></extra>",
                       marker=list(color=~colr))
gfig

```