---
title: "transmission-v-batteries"
author: "GeoffRussell"
date: "2024-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
comma<-function(x) prettyNum(signif(x,digits=4),big.mark=",")
options(scipen=100000)
SAcust<-919000
```

## Introduction

If you have a city needing power and a coal mine 200kms away, it has long been true that building the power plant at the mine and transmitting
the electricity by cable to the city is cheaper than building the generator at the city and shipping the coal. 

But what about the choice between transmission and batteries? Assuming you have excess generation in two locations, but at different times, then what's
the optimal way of avoiding curtailment of the power? Should you build an interconnector or install batteries in both locations?

## Transmission or batteries 

Let's start with a simple model.

We have locations $A$ and $B$ with local generators. $A$ and $B$ are separated by $k$ kilometers and we want to know whether it's best (cheapest) to
connect $A$ and $B$ with transmission $T$ or build batteries. 

We'll use South Australia and NSW as a concrete case study. Project *EnergyConnect* is currently under construction. 
It's about 900kms long and will cost some \$2.28 billion with a construction time
of [about 5 years](https://www.projectenergyconnect.com.au/); starting from the EIS approval in January 2022.

Initially, PEC was proposed *before* the 2016 Statewide blackout in SA in 2016. The cost range was originally estimated as between \$300m and \$700m. See [here](https://wattclarity.com.au/articles/2024/07/a-tale-of-two-mega-projects-project-energy-connect-and-snowy-2-0-timelines-part-1) for a chronology of the project and its costs.

```{r}
PECcap<-800e6
PECcost<-2.28e9
dperkm<-1.93e6
km<-900
cNZAPEC<-dperkm*km
diff<-comma((PECcost-cNZAPEC)/1e6)
```
The power capacity of PEC is `r comma(PECcap/1e6)` MW.

The Net Zero Australia [methods report](https://www.netzeroaustralia.net.au/wp-content/uploads/2023/04/Net-Zero-Australia-Methods-Assumptions-Scenarios-Sensitivities.pdf) assumes a cost of $1.93m/km for a twin 500V HVAC connector with a `r comma(PECcap/1e6)`MW capacity. PEC, at \$2.28b is considerably more expensive by some \$`r diff` million; which mades the NZA assumptions questionable. 

## The symmetry of PEC 

SA has `r comma(SAcust)` customers, far fewer than NSW. The relationship between the two sides of PEC isn't clear (to me).  SA has a considerable number of wind farms which
may wish to export electricity elsewhere. So as coal plants close in NSW, export flows from SA to NSW may rise. At present, however the flows on existing
interconnecters are are predominantly into SA.

## Simple battery replacement

Assuming we wanted to replace PEC with batteries, how many could we buy at both ends? Gencost 2024, Table B.6 puts the 12 hour battery cost in 2025 of \$381/kWh. 
This is roughly half the current retail cost per kWh of a household Tesla Powerwall 2. These hold 13.5 kWh and cost anywhere from \$10k to \$13k.

```{r}
bcost12<-381
kwhEnd<-PECcost/bcost12/2  # kWh
kwhEnd
```

We could put `r comma(kwhEnd/1e6)` GWh of batteries at each end. Given that PEC is a `r comma(PECcap/1e6)`MW connection. It will 
deliver `r comma(PECcap/1e6)`MWh of energy per hour.  How
does this compare with our batteries? If we were running them at full power `r comma(PECcap/1e6)`MW, how long would they last if full?

```{r}
bTime<-kwhEnd*1e3/PECcap
bTime
```
The batteries would supply some `r comma(bTime)` hours at full power: `r comma(PECcap/1e6)` MW.

## Replacement with home batteries

We could alternatively assume that people would buy home batteries as a substitute for PEC. How many batteries could you get for the cost of PEC? And
how long would they last? We assume, optimistically, they are being managed by the grid 
so that they can effectively supply a similar power to PEC for some given time. We could choose to put the batteries evenly in both states, but for now
we'll put them all in SA.



```{r}
SAfraction<-1
pwallCost<-c(10000,13000)
pwallE<-13.5e3
nPwall<-PECcost/pwallCost
timePwall<-nPwall*pwallE/PECcap
frPwall<-nPwall/(SAcust*SAfraction)
nPwall
timePwall
```

In this scenario we'll have between 
`r comma(nPwall[2])` and
`r comma(nPwall[1])` batteries (`r comma(frPwall[2]*100)`% to `r comma(frPwall[1]*100)`%) of SA customers.

The lifespan of PEC will be measured in decades. In contrast the batteries and associated electronics will need replacing periodically.
